---
title: "Climate emergency slide pack"
output: powerpoint_presentation
params:
  la: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error=TRUE, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8, fig.width=12)

library(tidyverse) ; library(sf) ; library(scales) ; library(sf) ; library(treemapify) ; library(ggsci) ; library(knitr)

la <- st_read(paste0("https://ons-inspire.esriuk.com/arcgis/rest/services/Administrative_Boundaries/Local_Authority_Districts_April_2019_Boundaries_UK_BGC/MapServer/0/query?where=UPPER(lad19cd)%20like%20'%25", URLencode(toupper(params$la), reserved = TRUE), "%25'&outFields=lad19cd,lad19nm,long,lat&outSR=4326&f=geojson"), quiet = TRUE) %>% 
  select(area_code = lad19cd, area_name = lad19nm, lon = long, lat)

id <- filter(lookup, area_code == params$la)$area_name
```

These slides (hopefully!) include visualisations of open data relating to climate change. The visualisations are licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/).

#### Carbon dioxide emissions
```{r}
co2_emissions <- co2_emissions %>% 
  filter(area_code == params$la,
         group %in% c("Domestic total", "Industry and commercial total", "Transport total")) 

co2_emissions_test = nrow(co2_emissions) != 0
```

```{r eval=co2_emissions_test}
ggplot(co2_emissions, aes(x = period, y = value, fill = group)) +
  geom_col(position = "stack") +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(labels = comma, expand = c(0.005, 0.005)) +
  scale_fill_viridis_d(direction = -1,
                       labels = c("Domestic total" = "Domestic", 
                                  "Industry and commercial total" = "Industry and commercial", 
                                  "Transport total" = "Transport")) +
  labs(x = NULL, 
       y = expression(paste(italic("kt"), CO[2])),
       title = "Carbon dioxide emissions by main sector",
       subtitle = paste0(id, ", 2005-2017"),
       caption = "Source: BEIS",
       fill = NULL) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_x()
```

#### Carbon dioxide emissions from large industrial installations
```{r}
large_point_sources <- large_point_sources %>% 
  mutate(value = round(value,0)) %>% 
  filter(area_code == params$la)

large_point_sources_test = nrow(large_point_sources) != 0
```

```{r eval=large_point_sources_test}
ggplot(large_point_sources, aes(lon, lat)) + 
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_point(aes(size = value, alpha = value), fill = "#212121", shape = 21) +
  scale_alpha_continuous(guide = "legend", label = comma, breaks = pretty_breaks(5)) +
  scale_size_continuous(label = comma, range = c(2, 12), breaks = pretty_breaks(5)) +
  labs(x = NULL, y = NULL, size = "Tonnes", alpha = "Tonnes",
       title = expression(bold(paste(CO[2], " emissions from large industrial installations"))),
       subtitle = paste0(id, ", 2016"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: NAEI") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Non domestic energy consumption
```{r}
non_domestic_energy_consumption <- non_domestic_energy_consumption %>% 
  filter(area_code == params$la) %>% 
  mutate(value = round(value,0)) %>% 
  arrange(value) %>% 
  mutate(group = factor(group, levels = group)) 

non_domestic_energy_consumption_test = nrow(non_domestic_energy_consumption) != 0
```


```{r eval=non_domestic_energy_consumption_test}
ggplot(non_domestic_energy_consumption, aes(x = value, y = group)) +
  geom_segment(aes(x = 0, xend = value, y = group, yend = group), 
               linetype = "dotted", colour = "#212121", size = 0.5) + 
  geom_point(colour = "#dd1c77", size = 15) +
  geom_text(aes(label = comma(value)), fontface = "bold", size = 4, colour = "white") +
  labs(x = "Gigawatt hours (GWh)", y = NULL, 
       title = "Industrial and commercial fuel use",
       subtitle = paste0(id, ", 2016"),
       caption = "Source: BEIS",
       fill = NULL) +
  theme_x() +
  theme(aspect.ratio = 0.5,
        panel.grid.major = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0))
```

#### Road transport fuel consumption
```{r}
road_transport_fuel_consumption <- road_transport_fuel_consumption %>% 
  filter(area_code == params$la)

road_transport_fuel_consumption_test = nrow(road_transport_fuel_consumption) != 0
```

```{r eval=road_transport_fuel_consumption_test}
ggplot(road_transport_fuel_consumption, aes(x = period, y = value, colour = group, group = group)) +
  geom_line(size = 0.8) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", labels = comma,
                     expand = c(0.005, 0.005)) +
  scale_color_futurama() +
  guides(colour = guide_legend(nrow = 1)) +
  labs(x = NULL, 
       y = "tonnes of oil",
       title = "Fuel use by vehicle type",
       subtitle = paste0(id, ", 2005-2017"),
       caption = "Source: BEIS",
       colour = NULL) +
  theme_x() + 
  theme(axis.title.y = element_text(hjust = 0))
```

#### Domestic energy consumption
```{r}
domestic_energy_consumption <- domestic_energy_consumption %>% 
  filter(area_code == params$la) %>% 
  mutate(value = round(value,0)) %>% 
  arrange(value) %>% 
  mutate(group = factor(group, levels = group)) 

domestic_energy_consumption_test = nrow(domestic_energy_consumption) != 0
```

```{r eval=domestic_energy_consumption_test}
ggplot(domestic_energy_consumption, aes(x = value, y = group)) +
  geom_segment(aes(x = 0, xend = value, y = group, yend = group), 
               linetype = "dotted", colour = "#212121", size = 0.5) + 
  geom_point(colour = "#DDCC77", size = 15) +
  geom_text(aes(label = comma(value)), fontface = "bold", size = 4, colour = "white") +
  labs(x = "Gigawatt hours (GWh)", y = NULL, 
       title = "Domestic fuel use",
       subtitle = paste0(id, ", 2016"),
       caption = "Source: BEIS",
       fill = NULL) +
  theme_x() +
  theme(aspect.ratio = 0.5,
        panel.grid.major = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0))
```

#### Domestic property build period
```{r}
domestic_property_build_period <- domestic_property_build_period %>% 
  filter(area_code == params$la) %>% 
  mutate(group = as.factor(group),
         group = fct_relevel(group, "pre-1900")) 

domestic_property_build_period_test = nrow(domestic_property_build_period) != 0
```


```{r eval=domestic_property_build_period_test}
ggplot(domestic_property_build_period, aes(x = factor(group), y = value)) +
  geom_col(fill = "brown4") +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, 
       y = "Dwellings",
       title = "Build period of domestic dwellings",
       subtitle = paste0(id, ", 2018"),
       caption = "Source: Valuation Office Agency")  +
  theme_x()
```

#### Renewable electricity generation
```{r}
renewable_electricity_generation <- renewable_electricity_generation %>% 
  filter(area_code == params$la, value != 0) %>%
  mutate(info = str_c(group, ": ",  comma(value)))

renewable_electricity_generation_test = nrow(renewable_electricity_generation) != 0
```

```{r eval=renewable_electricity_generation_test}
ggplot(renewable_electricity_generation, aes(area = value, fill = fct_reorder(group, value, .desc = TRUE), subgroup = group, label = info)) +
  geom_treemap(colour = "#212121") +
  geom_treemap_text(colour = "#FFFFFF", place = "bottomleft", reflow = TRUE, 
                    padding.x = grid::unit(1.5, "mm"),
                    padding.y = grid::unit(2, "mm"),
                    size = 14) +
  scale_fill_simpsons() +
  labs(x = NULL, y = NULL, 
       title = "Renewable electricity generation by source",
       subtitle = paste0(id, ", 2017"),
       caption = "Source: BEIS",
       fill = "MWh") +
  theme_x() +
  theme(legend.position = "right") 
```

#### Recycling
```{r}
recycling <- recycling %>% 
  filter(area_code == params$la)

recycling_test = nrow(recycling) != 0
```

```{r eval=recycling_test}
ggplot(recycling, aes(x = period, y = value, group = 1)) +
  geom_line(colour = "#117733", size = 1) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL,
       title = "Reuse, recycling or composting of household waste",
       subtitle = paste0(id, ", 2010/11 - 2017/18"),
       caption = "Source: DEFRA") +
  theme_x()
```

#### Licensed electric vehicles
```{r}
electric_vehicles <- electric_vehicles %>% 
  filter(area_code == params$la) 

electric_vehicles_test = nrow(electric_vehicles) != 0
```

```{r eval=electric_vehicles_test}
ggplot(electric_vehicles, aes(x = period, y = value)) +
  geom_line(colour = "#332288", size = 1.5) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", limits = c(0, NA), labels = comma) +
  labs(x = NULL, y = NULL, 
       title = "Number of licensed electric vehicles",
       subtitle = paste0(id, ", 2011 Q4 - 2019 Q2"),
       caption = "Data: DfT and DVLA") +
  theme_x() +
  theme(axis.title.y = element_text(hjust = 0))
```

#### Cycling
```{r}
cycling <- cycling %>% 
  filter(area_code == params$la) 

cycling_test = nrow(cycling) != 0
```

```{r eval=cycling_test}
cycling %>% 
  mutate(group = paste("At least", tolower(group)),
         value = round(value,1)) %>% 
    select(`How often?` = group, `%` = value) %>% 
  kable(caption = "Proportion of residents who cycle, 2017/18 (Source: Department for Transport)")
```

#### Walking
```{r}
walking <- walking %>% 
  filter(area_code == params$la) 

walking_test = nrow(walking) != 0
```

```{r eval=walking_test}
walking %>% 
  mutate(group = paste("At least", tolower(group)),
         value = round(value,1)) %>% 
    select(`How often?` = group, `%` = value) %>% 
  kable(caption = "Proportion of residents who walk, 2017/18 (Source: Department for Transport)")
```