---
title: '`r paste("Climate emergency slide pack for ", filter(lookup, area_code == params$la)$area_name)`'
output: powerpoint_presentation
params:
  la: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error=TRUE, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8, fig.width=12)

library(tidyverse) ; library(sf) ; library(scales) ; library(sf) ; library(treemapify) ; library(ggsci) ; library(knitr)

la <- filter(la, area_code == params$la)
lsoa <- filter(lsoa, area_code == params$la)

id <- filter(lookup, area_code == params$la)$area_name
```

These slides (hopefully!) include visualisations of open data relating to climate change. The visualisations are licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/). You can find the data behind these visualisations in our [open data companion](http://www.trafforddatalab.io/climate_emergency/).

#### Carbon dioxide emissions
```{r}
co2_emissions <- co2_emissions %>% 
  filter(area_code == params$la,
         group %in% c("Domestic total", "Industry and commercial total", "Transport total")) 

co2_emissions_test = nrow(co2_emissions) != 0
```

```{r eval=co2_emissions_test}
ggplot(co2_emissions, aes(x = period, y = value, fill = group)) +
  geom_col(position = "stack") +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(labels = comma, expand = c(0.005, 0.005)) +
  scale_fill_viridis_d(direction = -1,
                       labels = c("Domestic total" = "Domestic", 
                                  "Industry and commercial total" = "Industry and commercial", 
                                  "Transport total" = "Transport")) +
  labs(x = NULL, 
       y = expression(paste(italic("kt"), CO[2])),
       title = "Carbon dioxide emissions by main sector",
       subtitle = paste0(id, ", 2005-2017"),
       caption = "Source: BEIS",
       fill = NULL) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_x()
```

#### Carbon dioxide emissions from large industrial installations
```{r}
large_point_sources <- large_point_sources %>% 
  mutate(value = round(value,0)) %>% 
  filter(area_code == params$la)

large_point_sources_test = nrow(large_point_sources) != 0
```

```{r eval=large_point_sources_test}
ggplot(large_point_sources, aes(lon, lat)) + 
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_point(aes(size = value, alpha = value), fill = "#212121", shape = 21) +
  scale_alpha_continuous(guide = "legend", label = comma, breaks = pretty_breaks(5)) +
  scale_size_continuous(label = comma, range = c(2, 12), breaks = pretty_breaks(5)) +
  labs(x = NULL, y = NULL, size = "Tonnes", alpha = "Tonnes",
       title = expression(bold(paste(CO[2], " emissions from large industrial installations"))),
       subtitle = paste0(id, ", 2016"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: NAEI") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Non domestic energy consumption
```{r}
non_domestic_energy_consumption <- non_domestic_energy_consumption %>% 
  filter(area_code == params$la) %>% 
  mutate(value = round(value,0)) %>% 
  arrange(value) %>% 
  mutate(group = factor(group, levels = group)) 

non_domestic_energy_consumption_test = nrow(non_domestic_energy_consumption) != 0
```


```{r eval=non_domestic_energy_consumption_test}
ggplot(non_domestic_energy_consumption, aes(x = value, y = group)) +
  geom_segment(aes(x = 0, xend = value, y = group, yend = group), 
               linetype = "dotted", colour = "#212121", size = 0.5) + 
  geom_point(colour = "#dd1c77", size = 15) +
  geom_text(aes(label = comma(value)), fontface = "bold", size = 4, colour = "white") +
  labs(x = "Gigawatt hours (GWh)", y = NULL, 
       title = "Industrial and commercial fuel use",
       subtitle = paste0(id, ", 2016"),
       caption = "Source: BEIS",
       fill = NULL) +
  theme_x() +
  theme(aspect.ratio = 0.5,
        panel.grid.major = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0))
```

#### Road transport fuel consumption
```{r}
road_transport_fuel_consumption <- road_transport_fuel_consumption %>% 
  filter(area_code == params$la)

road_transport_fuel_consumption_test = nrow(road_transport_fuel_consumption) != 0
```

```{r eval=road_transport_fuel_consumption_test}
ggplot(road_transport_fuel_consumption, aes(x = period, y = value, colour = group, group = group)) +
  geom_line(size = 0.8) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", labels = comma,
                     expand = c(0.005, 0.005)) +
  scale_color_futurama() +
  guides(colour = guide_legend(nrow = 1)) +
  labs(x = NULL, 
       y = "tonnes of oil",
       title = "Fuel use by vehicle type",
       subtitle = paste0(id, ", 2005-2017"),
       caption = "Source: BEIS",
       colour = NULL) +
  theme_x() + 
  theme(axis.title.y = element_text(hjust = 0))
```

#### Domestic energy consumption
```{r}
domestic_energy_consumption <- domestic_energy_consumption %>% 
  filter(area_code == params$la) %>% 
  mutate(value = round(value,0)) %>% 
  arrange(value) %>% 
  mutate(group = factor(group, levels = group)) 

domestic_energy_consumption_test = nrow(domestic_energy_consumption) != 0
```

```{r eval=domestic_energy_consumption_test}
ggplot(domestic_energy_consumption, aes(x = value, y = group)) +
  geom_segment(aes(x = 0, xend = value, y = group, yend = group), 
               linetype = "dotted", colour = "#212121", size = 0.5) + 
  geom_point(colour = "#DDCC77", size = 15) +
  geom_text(aes(label = comma(value)), fontface = "bold", size = 4, colour = "white") +
  labs(x = "Gigawatt hours (GWh)", y = NULL, 
       title = "Domestic fuel use",
       subtitle = paste0(id, ", 2016"),
       caption = "Source: BEIS",
       fill = NULL) +
  theme_x() +
  theme(aspect.ratio = 0.5,
        panel.grid.major = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0))
```

#### Domestic property build period
```{r}
domestic_property_build_period <- domestic_property_build_period %>% 
  filter(area_code == params$la) %>% 
  mutate(group = as.factor(group),
         group = fct_relevel(group, "pre-1900")) 

domestic_property_build_period_test = nrow(domestic_property_build_period) != 0
```

```{r eval=domestic_property_build_period_test}
ggplot(domestic_property_build_period, aes(x = factor(group), y = value)) +
  geom_col(fill = "brown4") +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, 
       y = "Dwellings",
       title = "Build period of domestic dwellings",
       subtitle = paste0(id, ", 2018"),
       caption = "Source: Valuation Office Agency")  +
  theme_x()
```

#### Fuel poverty
```{r}
# Fuel poverty
fuel_poverty <- left_join(lsoa,
                          filter(fuel_poverty, area_code == params$la), by = "lsoa11cd")

fuel_poverty_test = nrow(fuel_poverty) != 0
```

```{r eval=fuel_poverty_test}
ggplot(fuel_poverty) + 
  geom_sf(aes(fill = value), color = "#FFFFFF", size = 0.5, alpha = 0.8) +
  scale_fill_viridis(discrete = F, 
                     label = function(x) paste0(x, "%"),
                     direction = -1,
                     guide = guide_colourbar(
                       direction = "vertical",
                       barwidth = unit(3, units = "mm"),
                       title.position = 'top',
                       title.vjust = 1)) +
  labs(x = NULL, y = NULL,
       title = "Proportion of households in fuel poverty",
       subtitle = paste0(id, ", 2017"),
              caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: BEIS",
       fill = NULL) +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.title = element_text(size = 16),
        legend.position = "right",
        legend.text = element_text(size = 6))
```

#### Background N02 concentrations
```{r}
# Background air pollution
background_air_pollution <- background_air_pollution %>% 
  filter(area_code == params$la)

background_air_pollution_test = nrow(background_air_pollution) != 0
```

```{r eval=background_air_pollution_test}
ggplot(data = background_air_pollution, 
       aes(fill = NO2, color = NO2)) +
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_sf(color = NA, alpha = 0.8) +
  scale_fill_viridis_c(option = "magma", direction = -1,
                       guide = guide_colorbar(
                         barheight = unit(2, units = "mm"),
                         barwidth = unit(50, units = "mm"),
                         title.position = "left",
                         title.vjust = 1,
                         label.hjust = 0.5
                       )) +
  labs(title = expression(bold(paste("Modelled background ", NO[2], " concentrations"))),
       subtitle = paste0(id, ", 2017"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: DEFRA",
       fill = "Annual mean (μg/m3)") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Background PM10 concentrations
```{r eval=background_air_pollution_test}
ggplot(data = background_air_pollution, 
       aes(fill = PM10, color = PM10)) +
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_sf(color = NA, alpha = 0.8) +
  scale_fill_viridis_c(option = "magma", direction = -1,
                       guide = guide_colorbar(
                         barheight = unit(2, units = "mm"),
                         barwidth = unit(50, units = "mm"),
                         title.position = "left",
                         title.vjust = 1,
                         label.hjust = 0.5
                       )) +
  labs(title = expression(bold(paste("Modelled background ", PM[10], " concentrations"))),
       subtitle = paste0(id, ", 2017"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: DEFRA",
       fill = "Annual mean (μg/m3)") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Background PM2.5 concentrations
```{r eval=background_air_pollution_test}
ggplot(data = background_air_pollution, 
       aes(fill = PM2_5, color = PM2_5)) +
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_sf(color = NA, alpha = 0.8) +
  scale_fill_viridis_c(option = "magma", direction = -1,
                       guide = guide_colorbar(
                         barheight = unit(2, units = "mm"),
                         barwidth = unit(50, units = "mm"),
                         title.position = "left",
                         title.vjust = 1,
                         label.hjust = 0.5
                       )) +
  labs(title = expression(bold(paste("Modelled background ", PM[2.5], " concentrations"))),
       subtitle = paste0(id, ", 2017"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: DEFRA",
       fill = "Annual mean (μg/m3)") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Renewable electricity generation
```{r}
renewable_electricity_generation <- renewable_electricity_generation %>% 
  filter(area_code == params$la, value != 0) %>%
  mutate(info = str_c(group, ": ",  comma(value)))

renewable_electricity_generation_test = nrow(renewable_electricity_generation) != 0
```

```{r eval=renewable_electricity_generation_test}
ggplot(renewable_electricity_generation, aes(area = value, fill = fct_reorder(group, value, .desc = TRUE), subgroup = group, label = info)) +
  geom_treemap(colour = "#212121") +
  geom_treemap_text(colour = "#FFFFFF", place = "bottomleft", reflow = TRUE, 
                    padding.x = grid::unit(1.5, "mm"),
                    padding.y = grid::unit(2, "mm"),
                    size = 14) +
  scale_fill_simpsons() +
  labs(x = NULL, y = NULL, 
       title = "Renewable electricity generation by source",
       subtitle = paste0(id, ", 2017"),
       caption = "Source: BEIS",
       fill = "MWh") +
  theme_x() +
  theme(legend.position = "right") 
```

#### Onshore wind
```{r}
onshore_wind <- onshore_wind %>% 
  filter(area_code == params$la)

onshore_wind_test = nrow(onshore_wind) != 0
```

```{r eval=onshore_wind_test}
ggplot(onshore_wind, aes(lon, lat)) + 
  geom_sf(data = la, fill = NA, color = "#212121", size = 0.5) +
  geom_point(color = "#756bb1", size = 3) +
  labs(x = NULL, y = NULL,
       title = "Onshore wind turbines",
       subtitle = paste0(distinct(onshore_wind, area_name), " 2019-06"),
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019\nSource: BEIS") +
  coord_sf(datum = NA) +
  theme_x() +
  theme(plot.caption = element_text(size = 8, hjust = 0),
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)) 
```

#### Recycling
```{r}
recycling <- recycling %>% 
  filter(area_code == params$la)

recycling_test = nrow(recycling) != 0
```

```{r eval=recycling_test}
ggplot(recycling, aes(x = period, y = value, group = 1)) +
  geom_line(colour = "#117733", size = 1) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL,
       title = "Reuse, recycling or composting of household waste",
       subtitle = paste0(id, ", 2010/11 - 2017/18"),
       caption = "Source: DEFRA") +
  theme_x()
```

#### Licensed electric vehicles
```{r}
electric_vehicles <- electric_vehicles %>% 
  filter(area_code == params$la) 

electric_vehicles_test = nrow(electric_vehicles) != 0
```

```{r eval=electric_vehicles_test}
ggplot(electric_vehicles, aes(x = period, y = value)) +
  geom_line(colour = "#332288", size = 1.5) +
  geom_hline(yintercept = 0, size = 1, colour = "#212121") +
  scale_y_continuous(position = "right", limits = c(0, NA), labels = comma) +
  labs(x = NULL, y = NULL, 
       title = "Number of licensed electric vehicles",
       subtitle = paste0(id, ", 2011 Q4 - 2019 Q2"),
       caption = "Data: DfT and DVLA") +
  theme_x() +
  theme(axis.title.y = element_text(hjust = 0))
```

#### Cycling
```{r}
cycling <- cycling %>% 
  filter(area_code == params$la) 

cycling_test = nrow(cycling) != 0
```

```{r eval=cycling_test}
cycling %>% 
  mutate(group = paste("At least", tolower(group)),
         value = round(value,1)) %>% 
    select(`How often?` = group, `%` = value) %>% 
  kable(caption = "Proportion of residents who cycle, 2017/18. Source: Department for Transport")
```

#### Walking
```{r}
walking <- walking %>% 
  filter(area_code == params$la) 

walking_test = nrow(walking) != 0
```

```{r eval=walking_test}
walking %>% 
  mutate(group = paste("At least", tolower(group)),
         value = round(value,1)) %>% 
    select(`How often?` = group, `%` = value) %>% 
  kable(caption = "Proportion of residents who walk, 2017/18. Source: Department for Transport")
```

#### Green belt
```{r}
green_belt <- green_belt %>% 
  filter(area_code == params$la) 

green_belt_test = nrow(green_belt) != 0
```

```{r eval=green_belt_test, results='asis'}
cat(paste("The extent of the designated Green Belt in Trafford in 2019 was estimated at ", comma(pull(filter(green_belt,area_name == "Trafford"), green_belt_area)), " hectares, or approximately ", pull(summarise(filter(green_belt, area_name == "Trafford"), percent = round(green_belt_area/total_area*100, 1)),percent), "% of the land area of Trafford."))
```

